---
title: "Complete Analysis - Effect of Economic Crisis"
author: "Sharmila Biswas"
output: html_document
---

Github link: https://github.com/Shormi5399/Data-Science-I-Project

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(dplyr)
library(forecast)
library(lmtest)
```

## Exploratory Analysis

How many unique observations to you have?

The number of unique observations for CPI, GDP, unemployment, manufacturing and trade respectively are:

```{r}
economy_data <- read.csv("economy_data.csv")
nrow(economy_data[!is.na(economy_data$CPI),])
nrow(economy_data[!is.na(economy_data$GDP),])
nrow(economy_data[!is.na(economy_data$Unemployment.Rate),])
nrow(economy_data[!is.na(economy_data$Manufacturing.Rate),])
nrow(economy_data[!is.na(economy_data$Trade),])
```

What information/features/characteristics do you have for each observation?

We modify the scope of our dataset from 1991 to 2014, since unemployment data is only available for this range.

```{r}
economy_data <- economy_data[economy_data$Year>=1991 & economy_data$Year<=2014,]
head(economy_data)
```

What are the min/max/mean/median/sd values for each of these features?

```{r summary_stats_cpi}
#Summary values for CPI
summary(economy_data$CPI,na.rm = TRUE)
sd(economy_data$CPI, na.rm = TRUE)

#Summary values for GDP
summary(economy_data$GDP,na.rm = TRUE)
sd(economy_data$GDP, na.rm = TRUE)

#Summary values for Unemployment
summary(economy_data$Unemployment.Rate,na.rm = TRUE)
sd(economy_data$Unemployment.Rate, na.rm = TRUE)

#Summary values for Manufacturing
summary(economy_data$Manufacturing.Rate,na.rm = TRUE)
sd(economy_data$Manufacturing.Rate, na.rm = TRUE)

#Summary values for Trade
summary(economy_data$Trade,na.rm = TRUE)
sd(economy_data$Trade, na.rm = TRUE)

```

We found the averages for CPI, gdp, unemployment, manufacturing and trade for all the countries.
Now, let's manipulate the dataset to include a column to indicate if the time period for the data is before(1991-2006) or after(2007-2014) the economic crisis.


```{r}
economy_data <- economy_data %>% mutate(Time_Period = ifelse(Year < 2007,"Before Crisis","After Crisis"))
head(economy_data)
```

Converting Time_Period to a factor variable and releveling the labels.

```{r}
economy_data$Time_Period <- as.character(economy_data$Time_Period)
economy_data$Time_Period <- factor(economy_data$Time_Period, levels = c("Before Crisis", "After Crisis"))
levels(economy_data$Time_Period)
```


```{r}
ggplot(economy_data, aes(x=Time_Period, y = Unemployment.Rate))+geom_boxplot(na.rm = T)
```

The boxplot confirms the finding the median unemployment is almost the same before and after crisis.
This is an interesting finding since the 2007 economic crisis led to unemployment. We will try to analyze how the unemployment trends varied post recession by drilling down further.


```{r}
ggplot(economy_data, aes(x=Time_Period, y = CPI))+geom_boxplot(na.rm = T)
```

The median CPI has increased increased post the crisis but the variability has got much less over time.

```{r}
ggplot(economy_data, aes(x=Time_Period, y = GDP))+ylim(0,100000000)+geom_boxplot(na.rm = T)
```

The median gdp has decreased post the crisis but the variability has got much less over time.

```{r}
ggplot(economy_data, aes(x=Time_Period, y = Manufacturing.Rate))+geom_boxplot(na.rm = T)
```

The median manufacturing rate has decreased post the crisis.

```{r}
ggplot(economy_data, aes(x=Time_Period, y = Trade))+geom_boxplot(na.rm = T)
```


What is the distribution of the core features (show a histogram)?

Below are the histograms of the distribution of the CPI, GDP, Unemployment, Manufacturing and Trade before and after 2007 economic crisis.

```{r hist_cpi}
ggplot(economy_data, aes(x = CPI)) +
  geom_histogram(fill = "white",
                 color = "black",
                 binwidth = 5,na.rm = T)+facet_grid(.~Time_Period)
```

The distribution of CPI was significantly different before and after crisis. The count of the number of countries with higher inflation rates increased after crisis.

```{r}

ggplot(economy_data, aes(x = GDP)) +
  geom_histogram(fill = "white",
                 color = "black",na.rm = T)+facet_grid(.~Time_Period)

```


```{r}

ggplot(economy_data, aes(x = Unemployment.Rate)) +
  geom_histogram(fill = "white",
                 color = "black",
                 binwidth = 0.5,na.rm = T)+facet_grid(.~Time_Period)
```


```{r}

ggplot(economy_data, aes(x = Manufacturing.Rate)) +
  geom_histogram(fill = "white",
                 color = "black",na.rm = T)+facet_grid(.~Time_Period)

```

```{r}

ggplot(economy_data, aes(x = Trade)) +
  geom_histogram(fill = "white",
                 color = "black",na.rm = T)+facet_grid(.~Time_Period)

```

