---
title: "Complete Analysis - Effect of Economic Crisis"
author: "Sharmila Biswas"
output: html_document
---

Github link: https://github.com/Shormi5399/Data-Science-I-Project


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(dplyr)
library(forecast)
library(lmtest)
```

Let's understand the effect of CPI,GDP and Unemployment further by analyzing trend graph from 1991 to 2014 in North America(NAC),European Union(EUU) and South Asia(SAS).

Are there obvious trends in the data (over time, across subgroups, etc.), and are the differences statistically significant?

```{r}
region_data <- read.csv("region_data.csv")
unemp_compare <- region_data %>%
  select(Country.Name,Country.Code,Year,Unemployment.Rate,Time_Period) %>%
  filter(Year >= 1991 & Year<= 2014 & (Country.Code =='EUU'| Country.Code == 'SAS' | Country.Code == 'NAC')) %>%
  group_by(Year)

ggplot(unemp_compare,aes(x = Year, y = Unemployment.Rate, col = Country.Code, group = Country.Code)) +geom_line(na.rm = T,lwd = 1)+ theme(axis.text.x=element_text(angle=90, hjust=1))
```

There is a steep increase in the uemployment rates of North America and Europe while South Asia is comparitively stable post the 2008 recession.


```{r}
manufacturing_compare <- region_data %>%
  select(Country.Name,Country.Code,Year,Manufacturing.Rate) %>%
  filter(Year >= 1991 & Year<= 2014 & (Country.Code =='EUU'| Country.Code == 'SAS' | Country.Code == 'NAC')) %>%
  group_by(Year)

ggplot(manufacturing_compare,aes(x = Year, y = Manufacturing.Rate, col = Country.Code, group = Country.Code)) +geom_line(na.rm = T,lwd = 1)+ theme(axis.text.x=element_text(angle=90, hjust=1))
```


```{r}
trade_compare <- region_data %>%
  select(Country.Name,Country.Code,Year,Trade) %>%
  filter(Year >= 1991 & Year<= 2014 & (Country.Code =='EUU'| Country.Code == 'SAS' | Country.Code == 'NAC')) %>%
  group_by(Year)

ggplot(trade_compare,aes(x = Year, y = Trade, col = Country.Code, group = Country.Code)) +geom_line(na.rm = T,lwd = 1)+ theme(axis.text.x=element_text(angle=90, hjust=1))

```


```{r}
gdp_compare <- region_data %>%
  select(Country.Name,Country.Code,Year,GDP) %>%
  filter(Year >= 1991 & Year<= 2014 & (Country.Code =='EUU'| Country.Code == 'SAS' | Country.Code == 'NAC')) %>%
  group_by(Year)

ggplot(gdp_compare,aes(x = Year, y = GDP, col = Country.Code, group = Country.Code)) +geom_line(na.rm = T,lwd = 1)+ theme(axis.text.x=element_text(angle=90, hjust=1))

```

Understanding time series implementation for CPI in USA:

```{r}
CPI_trend<-read.csv("CPI_World_Bank.csv")
#plot(CPI_trend$)
colnames(CPI_trend)
#goal here is time period on x and CPI on y.
trends<-read.csv("Timeseries.csv")
class(trends)
#summary(lm(Time ~ CPI_USA, data=trends))
#plot(trends)
#abline(lsfit(x=trends$Time,y=trends$CPI_USA),col="red")
#install.packages('tseries') 
require(tseries) 
#install.packages('xts')
require(xts)
#converting data into time series data
trends.ts<-ts(as.vector(trends), start = c(1960,1), end = c(2015),frequency = 1)
trends.ts
class(trends.ts)
plot(trends.ts)
#Dickey-fuller test to see if the time series is stationary.
adf.test(trends.ts)
#Checking for p and q values through auto correlation function and partial auto correlation function.
acf(trends.ts)
pacf(trends.ts)
#this is an AR model
#install.packages('forecast')
auto.arima(trends.ts)
trends.arimausa<- arima(trends.ts, order=c(1,2,1))
trends.arimausa
trend.preds<-forecast.Arima(trends.arimausa,h=12)
plot.forecast(trend.preds)
summary(trend.preds)

#Hypothesis testing for USA CPI data
#install.packages('lmtest')
coeftest(trends.arimausa)
#z value is 2.2014, p=0.02771
#another way to obtain p value to verify
(1-pnorm(abs(trends.arimausa$coef)/sqrt(diag(trends.arimausa$var.coef))))*2

trends.resid <- residuals(trends.arimausa)
Box.test(trends.resid, lag = 10, type = "Ljung-Box")
```

Attempt to find trends in the CPI data.
This data set contains CPI for 265 countries over a period of 56 years.
The variable CPI is time sensitive.
We have used Dickey-fuller test to see if the time series is stationary and checked for p and q values through auto correlation function and partial auto correlation function.

p value obtained from this analysis is 0.027. Since CPI was an AR series, we will analyze this further on the statistical significance and its effect on the hypothesis.

What are the other salient aspects of the data (e.g. geospatial factors, text content, etc.)

The following plots are forecasts for each of the selected country's future unemployment rates based on the 1991-2014 data.  This also provides a projection of what the unemployment rates might have looked like in the absence of the data post the financial crisis

```{r} 
#NAC unemployment forecast vs actual, 2000-2007
par(mfrow=c(1,3))
trends.usa <-region_data %>% select(Year,Unemployment.Rate,Country.Code) %>% filter(Country.Code == "NAC")  
trendsts.usa <- ts(as.vector(trends.usa$Unemployment.Rate),start = c(1991,1), end = c(2014), frequency = 1)


trends.usabc <- region_data  %>% select(Year,Unemployment.Rate,Country.Code) %>% filter(Country.Code == "NAC" & Year >=1991 & Year<=2007)
trendsts.usabc <- ts(as.vector(trends.usabc$Unemployment.Rate),start = c(1991,1), end = c(2007), frequency = 1)
#install.packages("forecast")


trends.usaac <- region_data  %>% select(Year,Unemployment.Rate,Country.Code) %>% filter(Country.Code == "NAC" & Year >=1991 & Year<=2008)
trendsts.usaac <- ts(as.vector(trends.usabc$Unemployment.Rate),start = c(1991,1), end = c(2008), frequency = 1)
#install.packages("forecast")


trends.arimausa<- auto.arima(trendsts.usabc, seasonal = FALSE)
trends.arimausa
trend.preds<-forecast.Arima(trends.arimausa,h=12)
plot.forecast(trend.preds, ylim = c(2,10), xlim = c(2000,2010),xlab="NAC Unemployment fcst, 1991-2007 data", ylab="% Unemployment")

summary(trend.preds)
plot(trendsts.usa, ylim = c(2,10), xlim = c(2000,2010), xlab="NAC Actual Unemployment" ,ylab="% Unemployment")


trends.arimausa<- auto.arima(trendsts.usaac, seasonal = FALSE)
trends.arimausa
trend.preds<-forecast.Arima(trends.arimausa,h=12)
plot.forecast(trend.preds, ylim = c(2,10), xlim = c(2000,2010),xlab="NAC Unemployment fcst, 1991-2008 data", ylab="% Unemployment")
```


```{r}
par(mfrow=c(1,3))
#EU unemployment forecast vs actual, 2000-2007
#needs work
trends.eu <-region_data %>% select(Year,Unemployment.Rate,Country.Code) %>% filter(Country.Code == "EUU")  
trendsts.eu <- ts(as.vector(trends.eu$Unemployment.Rate),start = c(1991,1), end = c(2014), frequency = 1)

trends.eubc <- region_data  %>% select(Year,Unemployment.Rate,Country.Code) %>% filter(Country.Code == "EUU" & Year >=1991 & Year<=2007)
trendsts.eubc <- ts(as.vector(trends.eubc$Unemployment.Rate),start = c(1991,1), end = c(2007), frequency = 1)

trends.euac <- region_data  %>% select(Year,Unemployment.Rate,Country.Code) %>% filter(Country.Code == "EUU" & Year >=1991 & Year<=2008)
trendsts.euac <- ts(as.vector(trends.euac$Unemployment.Rate),start = c(1991,1), end = c(2008), frequency = 1)

trends.arimaeu<- auto.arima(trendsts.eubc, seasonal = FALSE) 
trends.arimaeu
trend.preds<-forecast.Arima(trends.arimaeu,h=12)
plot.forecast(trend.preds, ylim = c(2,10), xlim = c(2000,2010),xlab="EUU unemployment forecast, 1991-2007 data",ylab="% Unemployment")

summary(trend.preds)
plot(trendsts.eu, ylim = c(2,10), xlim = c(2000,2010), xlab="EUU Actual Unemployment",ylab="% Unemployment")

trends.arimaeu<- auto.arima(trendsts.euac,seasonal = FALSE)
trends.arimaeu
trend.preds<-forecast.Arima(trends.arimaeu,h=12)
plot.forecast(trend.preds, ylim = c(2,10), xlim = c(2000,2010),xlab="EUU Unemployment fcst, 1991-2008 data", ylab="% Unemployment")
```

```{r}
par(mfrow=c(1,3))
#SAS unemployment forecast vs actual, 2000-2007
trends.sas <-region_data %>% select(Year,Unemployment.Rate,Country.Code) %>% filter(Country.Code == "SAS")
trendsts.sas <- ts(as.vector(trends.sas$Unemployment.Rate),start = c(1991,1), end = c(2014), frequency = 1)

trends.sasbc <- region_data  %>% select(Year,Unemployment.Rate,Country.Code) %>% filter(Country.Code == "SAS" & Year >=1991 & Year<=2007)
trendsts.sasbc <- ts(as.vector(trends.sasbc$Unemployment.Rate),start = c(1991,1), end = c(2007), frequency = 1)

trends.sasac <- region_data  %>% select(Year,Unemployment.Rate,Country.Code) %>% filter(Country.Code == "SAS" & Year >=1991 & Year<=2008)
trendsts.sasac <- ts(as.vector(trends.sasac$Unemployment.Rate),start = c(1991,1), end = c(2008), frequency = 1)

trends.arimasas<- auto.arima(trendsts.sasbc) 
trend.preds<-forecast.Arima(trends.arimasas,h=12)
plot.forecast(trend.preds, ylim = c(2,10), xlim = c(2000,2010),xlab="SAS unemployment forecast, 1991-2007 data",ylab="% Unemployment")

summary(trend.preds)
plot(trendsts.sas, ylim = c(2,10), xlim = c(2000,2010), xlab="SAS Actual Unemployment",ylab="% unemployment")

trends.arimasas<- auto.arima(trendsts.sasac) 
trend.preds<-forecast.Arima(trends.arimasas,h=12)
plot.forecast(trend.preds, ylim = c(2,10), xlim = c(2000,2010),xlab="SAS unemployment forecast, 1991-2008 data",ylab="% Unemployment")
```

```{r}
#manufacturing rate US region
par(mfrow=c(1,3))
trends.usa <-region_data %>% select(Year,Manufacturing.Rate,Country.Code) %>% filter(Country.Code == "NAC")  
trendsts.usa <- ts(as.vector(trends.usa$Manufacturing.Rate),start = c(1991,1), end = c(2014), frequency = 1)

trends.usabc <- region_data  %>% select(Year,Manufacturing.Rate,Country.Code) %>% filter(Country.Code == "NAC" & Year >=1991 & Year<=2007)
trendsts.usabc <- ts(as.vector(trends.usabc$Manufacturing.Rate),start = c(1991,1), end = c(2007), frequency = 1)

trends.usaac <- region_data  %>% select(Year,Manufacturing.Rate,Country.Code) %>% filter(Country.Code == "NAC" & Year >=1991 & Year<=2008)
trendsts.usaac <- ts(as.vector(trends.usaac$Manufacturing.Rate),start = c(1991,1), end = c(2008), frequency = 1)

trends.arimausa<- auto.arima(trendsts.usabc)
trend.preds<-forecast.Arima(trends.arimausa,h=12)
plot.forecast(trend.preds, ylim = c(10,18), xlim = c(2000,2010), xlab="NAC mfg forecast, 1991-2007 data", ylab="rate")

summary(trend.preds)
plot(trendsts.usa, ylim = c(10,18), xlim = c(2000,2010), xlab="NAC Actual Mfg",ylab="% unemployment")

trends.arimausa<- auto.arima(trendsts.usaac)
trend.preds<-forecast.Arima(trends.arimausa,h=12)
plot.forecast(trend.preds, ylim = c(10,18), xlim = c(2000,2010), xlab="NAC mfg forecast, 1991-2008 data", ylab="rate")
```

```{r}
par(mfrow=c(1,3))
#manufacturing rate EUU region
trends.euu <-region_data %>% select(Year,Manufacturing.Rate,Country.Code) %>% filter(Country.Code == "EUU")
trendsts.euu <- ts(as.vector(trends.euu$Manufacturing.Rate),start = c(1991,1), end = c(2014), frequency = 1)

trends.eubc <- region_data  %>% select(Year,Manufacturing.Rate,Country.Code) %>% filter(Country.Code == "EUU" & Year >=1991 & Year<=2007)
trendsts.eubc <- ts(as.vector(trends.eubc$Manufacturing.Rate),start = c(1991,1), end = c(2007), frequency = 1)

trends.euac <- region_data  %>% select(Year,Manufacturing.Rate,Country.Code) %>% filter(Country.Code == "EUU" & Year >=1991 & Year<=2008)
trendsts.euac <- ts(as.vector(trends.euac$Manufacturing.Rate),start = c(1991,1), end = c(2008), frequency = 1)

trends.arimaeu<- auto.arima(trendsts.eubc)
trend.preds<-forecast.Arima(trends.arimaeu,h=12)
plot.forecast(trend.preds, ylim = c(12,20), xlim = c(2000,2010), xlab="EUU mfg forecast, 1991-2007 data", ylab="rate")

plot(trendsts.euu,ylim = c(12,20), xlim = c(2000,2010),xlab="EUU mfg rate", ylab="rate")

trends.arimaeu<- auto.arima(trendsts.euac)
trend.preds<-forecast.Arima(trends.arimaeu,h=12)
plot.forecast(trend.preds, ylim = c(12,20),xlim = c(2000,2010), xlab="EUU mfg forecast, 1991-2008 data", ylab="rate")
```

```{r}
par(mfrow=c(1,3))
#manufacturing rate SAS region
trends.sas <-region_data %>% select(Year,Manufacturing.Rate,Country.Code) %>% filter(Country.Code == "SAS")
trendsts.sas <- ts(as.vector(trends.sas$Manufacturing.Rate),start = c(1991,1), end = c(2014), frequency = 1)

trends.sasbc <- region_data  %>% select(Year,Manufacturing.Rate,Country.Code) %>% filter(Country.Code == "SAS" & Year >=1991 & Year<=2007)
trendsts.sasbc <- ts(as.vector(trends.sasbc$Manufacturing.Rate),start = c(1991,1), end = c(2007), frequency = 1)

trends.sasac <- region_data  %>% select(Year,Manufacturing.Rate,Country.Code) %>% filter(Country.Code == "SAS" & Year >=1991 & Year<=2008)
trendsts.sasac <- ts(as.vector(trends.sasac$Manufacturing.Rate),start = c(1991,1), end = c(2008), frequency = 1)

trends.arimasas<- auto.arima(trendsts.sasbc)
trend.preds<-forecast.Arima(trends.arimasas,h=12)
plot.forecast(trend.preds, ylim = c(14,20), xlim = c(2000,2010), xlab="SAS mfg forecast, 1991-2007 data", ylab="rate")

plot(trendsts.sas,ylim = c(14,20),xlim = c(2000,2010),xlab="SAS mfg rate", ylab="rate")

trends.arimasas<- auto.arima(trendsts.sasac)
trend.preds<-forecast.Arima(trends.arimasas,h=12)
plot.forecast(trend.preds,  ylim = c(14,20),xlim = c(2000,2010), xlab="SAS mfg forecast, 1991-2008 data", ylab="rate")
```

```{r}
par(mfrow=c(1,3))
#US trade
trends.usa <-region_data %>% select(Year,Trade,Country.Code) %>% filter(Country.Code == "NAC")  
trendsts.usa <- ts(as.vector(trends.usa$Trade),start = c(1991,1), end = c(2014), frequency = 1)

trends.usabc <- region_data  %>% select(Year,Trade,Country.Code) %>% filter(Country.Code == "NAC" & Year >=1991 & Year<=2007)
trendsts.usabc <- ts(as.vector(trends.usabc$Trade),start = c(1991,1), end = c(2007), frequency = 1)

trends.usaac <- region_data  %>% select(Year,Trade,Country.Code) %>% filter(Country.Code == "NAC" & Year >=1991 & Year<=2008)
trendsts.usaac <- ts(as.vector(trends.usaac$Trade),start = c(1991,1), end = c(2008), frequency = 1)

trends.arimausa<- auto.arima(trendsts.usabc)
trend.preds<-forecast.Arima(trends.arimausa,h=12)
plot.forecast(trend.preds, ylim = c(25,35),xlim = c(2000,2010),xlab="NAC trade forecast, 1991-2007 data", ylab="rate")

plot(trendsts.usa,  ylim = c(25,35),xlim = c(2000,2010),xlab="NAC actual trade rate", ylab="rate")

trends.arimausa<- auto.arima(trendsts.usaac)
trend.preds<-forecast.Arima(trends.arimausa,h=12)
plot.forecast(trend.preds,ylim = c(25,35), xlim = c(2000,2010),xlab="NAC trade forecast, 1991-2008 data", ylab="rate")
```

```{r}
par(mfrow=c(1,3))
#EU trade
trends.eu <-region_data %>% select(Year,Trade,Country.Code) %>% filter(Country.Code == "EUU")  
trendsts.eu <- ts(as.vector(trends.eu$Trade),start = c(1991,1), end = c(2014), frequency = 1)

trends.eubc <- region_data  %>% select(Year,Trade,Country.Code) %>% filter(Country.Code == "EUU" & Year >=1991 & Year<=2007)
trendsts.eubc <- ts(as.vector(trends.eubc$Trade),start = c(1991,1), end = c(2007), frequency = 1)

trends.euac <- region_data  %>% select(Year,Trade,Country.Code) %>% filter(Country.Code == "EUU" & Year >=1991 & Year<=2008)
trendsts.euac <- ts(as.vector(trends.euac$Trade),start = c(1991,1), end = c(2008), frequency = 1)

trends.arimaeu<- auto.arima(trendsts.eubc)
trend.preds<-forecast.Arima(trends.arimaeu,h=12)
plot.forecast(trend.preds,ylim = c(60,90), xlim = c(2000,2010),xlab="EUU trade forecast, 1991-2007 data", ylab="rate")

plot(trendsts.eu, ylim = c(60,90), xlim = c(2000,2010),xlab="EUU actual trade rate", ylab="rate")

trends.arimaeu<- auto.arima(trendsts.euac)
trend.preds<-forecast.Arima(trends.arimaeu,h=12)
plot.forecast(trend.preds, ylim = c(60,90),xlim = c(2000,2010),xlab="EUU trade forecast, 1991-2008 data", ylab="rate")
```

```{r}
par(mfrow=c(1,3))
#SAS trade
trends.sas <-region_data %>% select(Year,Trade,Country.Code) %>% filter(Country.Code == "SAS")  
trendsts.sas <- ts(as.vector(trends.sas$Trade),start = c(1991,1), end = c(2014), frequency = 1)

trends.sasbc <- region_data  %>% select(Year,Trade,Country.Code) %>% filter(Country.Code == "SAS" & Year >=1991 & Year<=2007)
trendsts.sasbc <- ts(as.vector(trends.sasbc$Trade),start = c(1991,1), end = c(2007), frequency = 1)

trends.sasac <- region_data  %>% select(Year,Trade,Country.Code) %>% filter(Country.Code == "SAS" & Year >=1991 & Year<=2008)
trendsts.sasac <- ts(as.vector(trends.sasac$Trade),start = c(1991,1), end = c(2008), frequency = 1)

trends.arimasas<- auto.arima(trendsts.sasbc)
trend.preds<-forecast.Arima(trends.arimasas,h=12)
plot.forecast(trend.preds,ylim = c(20,70), xlim = c(2000,2010),xlab="SAS trade forecast, 1991-2007 data", ylab="rate")

plot(trendsts.sas,ylim = c(20,70),xlim = c(2000,2010),xlab="SAS actual trade rate", ylab="rate")

trends.arimasas<- auto.arima(trendsts.sasac)
trend.preds<-forecast.Arima(trends.arimasas,h=12)
plot.forecast(trend.preds, ylim = c(20,70),xlim = c(2000,2010),xlab="SAS trade forecast, 1991-2008 data", ylab="rate")
```

```{r}
par(mfrow=c(1,3))
#US GDP
trends.usa <-region_data %>% select(Year,GDP,Country.Code) %>% filter(Country.Code == "NAC")  
trendsts.usa <- ts(as.vector(trends.usa$GDP),start = c(1991,1), end = c(2014), frequency = 1)

trends.usabc <- region_data  %>% select(Year,GDP,Country.Code) %>% filter(Country.Code == "NAC" & Year >=1991 & Year<=2007)
trendsts.usabc <- ts(as.vector(trends.usabc$GDP),start = c(1991,1), end = c(2007), frequency = 1)

trends.usaac <- region_data  %>% select(Year,GDP,Country.Code) %>% filter(Country.Code == "NAC" & Year >=1991 & Year<=2008)
trendsts.usaac <- ts(as.vector(trends.usaac$GDP),start = c(1991,1), end = c(2008), frequency = 1)

trends.arimausa<- auto.arima(trendsts.usabc)
trend.preds<-forecast.Arima(trends.arimausa,h=12)
plot.forecast(trend.preds,ylim = c(1.0e+13,2.0e+13), xlim = c(2000,2010), xlab="NAR GDP forecast, 1991-2007 data")

plot(trendsts.usa , ylim = c(1.0e+13,2.0e+13), xlim = c(2000,2010),xlab="NAR GDP", ylab ="point value")

trends.arimausa<- auto.arima(trendsts.usaac)
trend.preds<-forecast.Arima(trends.arimausa,h=12)
plot.forecast(trend.preds, ylim = c(1.0e+13,2.0e+13),xlim = c(2000,2010), xlab="NAR GDP forecast, 1991-2008 data")
```

```{r}
par(mfrow=c(1,3))
#EUU GDP
trends.eu <-region_data %>% select(Year,GDP,Country.Code) %>% filter(Country.Code == "EUU")  
trendsts.eu <- ts(as.vector(trends.eu$GDP),start = c(1991,1), end = c(2014), frequency = 1)

trends.eubc <- region_data  %>% select(Year,GDP,Country.Code) %>% filter(Country.Code == "EUU" & Year >=1991 & Year<=2007)
trendsts.eubc <- ts(as.vector(trends.eubc$GDP),start = c(1991,1), end = c(2007), frequency = 1)

trends.euac <- region_data  %>% select(Year,GDP,Country.Code) %>% filter(Country.Code == "EUU" & Year >=1991 & Year<=2008)
trendsts.euac <- ts(as.vector(trends.eubc$GDP),start = c(1991,1), end = c(2008), frequency = 1)

trends.arimaeu<- auto.arima(trendsts.eubc)
trend.preds<-forecast.Arima(trends.arimaeu,h=12)
plot.forecast(trend.preds, ylim=c(8e+12,2.5e+13),xlim = c(2000,2010), xlab="EU GDP forecast, 1991-2007 data")

plot(trendsts.eu , ylim=c(8e+12,2.5e+13), xlim = c(2000,2010),xlab="EU GDP")

trends.arimaeu<- auto.arima(trendsts.euac)
trend.preds<-forecast.Arima(trends.arimaeu,h=12)
plot.forecast(trend.preds, ylim=c(8e+12,2.5e+13),xlim = c(2000,2010), xlab="EU GDP forecast, 1991-2008 data")
```

```{r}
par(mfrow=c(1,3))
#SAS GDP
trends.sas <-region_data %>% select(Year,GDP,Country.Code) %>% filter(Country.Code == "SAS")  
trendsts.sas <- ts(as.vector(trends.sas$GDP),start = c(1991,1), end = c(2014), frequency = 1)

trends.sasbc <- region_data  %>% select(Year,GDP,Country.Code) %>% filter(Country.Code == "SAS" & Year >=1991 & Year<=2007)
trendsts.sasbc <- ts(as.vector(trends.sasbc$GDP),start = c(1991,1), end = c(2007), frequency = 1)

trends.sasac <- region_data  %>% select(Year,GDP,Country.Code) %>% filter(Country.Code == "SAS" & Year >=1991 & Year<=2008)
trendsts.sasac <- ts(as.vector(trends.sasbc$GDP),start = c(1991,1), end = c(2008), frequency = 1)

trends.arimasas<- auto.arima(trendsts.sasbc)
trend.preds<-forecast.Arima(trends.arimasas,h=12)
plot.forecast(trend.preds,ylim = c(0e+12, 3e+12),xlim = c(2000,2010), xlab="SAS GDP forecast, 1991-2007 data")

plot(trendsts.sas ,ylim = c(0e+12, 4e+12) ,xlim = c(2000,2010),xlab="SAS GDP")

trends.arimasas<- auto.arima(trendsts.sasac)
trend.preds<-forecast.Arima(trends.arimasas,h=12)
plot.forecast(trend.preds,ylim = c(0e+12, 3e+12),xlim = c(2000,2010), xlab="SAS GDP forecast, 1991-2008 data")
```

CPI data was unavailable at region-level. So we could not perform time series analysis for CPI.


